---
title: "Working with submission data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with submission data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`tally_submissions()` turns the Tally API's nested representation of a
form's submissions into a flat, wide tibble that's ready for the usual
data-wrangling tools. This vignette describes the shape of that tibble in
detail, then shows some common patterns for tidying it.

Since the code needs a live API key and a real form, it isn't run when
the vignette is built; the output shown is from an example form, a
conference feedback survey with a handful of questions.

```{r setup}
library(tallyr)
```

```{r, eval=FALSE}
feedback <- tally_submissions("3xLJ5V")
feedback
#> # A tibble: 42 × 6
#>   submission_id submitted_at        is_completed `What's your name?`
#>   <chr>         <dttm>              <lgl>        <chr>
#> 1 nWxyzA        2026-06-01 10:00:00 TRUE         Ada
#> 2 nWxyzB        2026-06-02 11:30:00 TRUE         Grace
#> 3 nWxyzC        2026-06-02 14:15:00 FALSE        NA
#> # ℹ 39 more rows
#> # ℹ 2 more variables: `Which sessions did you attend?` <chr>,
#> #   `Any other comments?` <chr>
```

## The shape of the tibble

There is one row per submission. The first three columns are always:

- `submission_id`: Tally's identifier for the submission.
- `submitted_at`: when it was submitted, as a POSIXct date-time in UTC.
- `is_completed`: whether the respondent finished the form, or abandoned
  it partway (a "partial" submission).

The remaining columns are the form's questions, in form order, one column
per question.

### Column names come from question titles

Each question column is named by the question's title, exactly as it
appears on the form -- including spaces and punctuation. That keeps the
columns self-describing, at the cost of names that need backticks in R
code:

```{r, eval=FALSE}
feedback$`What's your name?`
```

If the form has two questions with the same title, the names are made
unique by appending `...` and the column's position, in the same way as
`tibble::as_tibble()` repairs duplicate names:

```{r, eval=FALSE}
names(tally_submissions("9aKp3T"))
#> [1] "submission_id"        "submitted_at"         "is_completed"
#> [4] "What's your name?"    "Favourite colour?...5" "Favourite colour?...6"
```

### Answers are formatted strings

Every question column is a character vector, whatever the question type.
tallyr uses the formatted answer that the Tally API provides as a
consistent string representation across question types, falling back to
the raw answer (with multiple parts joined by `", "`) when there is no
formatted version.

In practice that means:

- Text questions give the text as entered.
- Multiple-choice and checkbox questions give the chosen option labels,
  comma-separated when more than one was selected (e.g.
  `"Teal, Purple"`).
- Numeric and date questions arrive as strings, so convert them yourself
  if you want to compute with them (see below).

A question a respondent didn't answer is `NA`. Partial submissions
typically have `NA` for every question after the point where the
respondent stopped.

## Tidying patterns

These examples use dplyr and friends, but nothing about the tibble
requires them -- base R works just as well.

### Renaming question columns

Question titles make awkward variable names, so renaming is usually the
first step:

```{r, eval=FALSE}
library(dplyr)

feedback <- tally_submissions("3xLJ5V") |>
  rename(
    name = `What's your name?`,
    sessions = `Which sessions did you attend?`,
    comments = `Any other comments?`
  )
```

### Converting answer types

Since answers arrive as strings, convert columns that are really numbers,
dates, or factors:

```{r, eval=FALSE}
ratings <- tally_submissions("5cRw8Y") |>
  rename(rating = `How would you rate the event? (1-5)`) |>
  mutate(rating = as.integer(rating))

ratings |>
  count(rating)
```

### Splitting multi-select answers

Checkbox answers with several selections come back as one
comma-separated string. To analyse the individual options, split them
into rows:

```{r, eval=FALSE}
library(tidyr)

feedback |>
  separate_longer_delim(sessions, delim = ", ") |>
  count(sessions, sort = TRUE)
```

### Working with submission times

`submitted_at` is a POSIXct in UTC. To report in your local time zone,
convert with `lubridate::with_tz()` (or `format()` in base R):

```{r, eval=FALSE}
feedback |>
  mutate(submitted_local = lubridate::with_tz(submitted_at, "Europe/London"))
```

## Filtering at import time

Some filtering is better done by the API than after import, since it
reduces what's downloaded. `tally_submissions()` can restrict by
completion status and by date:

```{r, eval=FALSE}
# only submissions where the respondent finished the form
tally_submissions("3xLJ5V", filter = "completed")

# only abandoned, partial submissions
tally_submissions("3xLJ5V", filter = "partial")

# a date window; accepts Dates, date-times, or ISO 8601 strings
tally_submissions("3xLJ5V", start_date = "2026-01-01", end_date = "2026-06-30")
```

Anything finer-grained than that -- filtering on answers, say -- is a job
for `dplyr::filter()` after import:

```{r, eval=FALSE}
feedback |>
  filter(!is.na(comments))
```

For an overview of the rest of the package, see `vignette("tallyr")`.
