---
title: "Troubleshooting your Tally setup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Troubleshooting your Tally setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

When tallyr isn't working, the cause is almost always one of a few
things: no API key where tallyr looks for it, a wrong or expired key, or
no internet connection. The package has two functions for telling these
apart: `tally_whoami()` as a quick yes/no check, and `tally_sitrep()`
for a full report.

The code in this vignette isn't run at build time; the output shown
illustrates what you'd see in each situation.

```{r setup}
library(tallyr)
```

## The quick check: tally_whoami()

`tally_whoami()` calls the Tally API and reports who you're
authenticated as. If it succeeds, your setup works:

```{r, eval = FALSE}
tally_whoami()
#> ✔ Authenticated with Tally as Ada Lovelace (ada@example.com)
```

If it errors, move on to `tally_sitrep()`.

## The full report: tally_sitrep()

`tally_sitrep()` checks everything relevant to tallyr working: which
accounts have keys available and which is active, whether the active
account's key was found (and where), whether it looks like a Tally key,
whether you're online, and whether the key actually authenticates. It
ends with recommendations for fixing anything that's wrong, and it never
errors, whatever state your setup is in.

A healthy setup looks like this:

```{r, eval=FALSE}
tally_sitrep()
#>
#> ── tallyr situation report ─────────────────────────────────────────────
#>
#> ── Accounts
#> ℹ Account with a key available: "default"
#> ℹ Active account: "default" (the default)
#>
#> ── API key
#> ✔ API key found
#> ℹ Source: Environment variable: TALLY_API_KEY
#> ✔ Key format looks valid (starts with "tly-")
#>
#> ── Connectivity
#> ✔ Internet: online
#> ✔ Authenticated as Ada Lovelace (ada@example.com)
#>
#> ── Recommendations
#> ✔ tallyr setup looks good!
```

The sections below walk through the situations it can catch.

## No API key found

```{r, eval = FALSE}
tally_sitrep()
#>
#> ── tallyr situation report ─────────────────────────────────────────────
#>
#> ── Accounts
#> ! No account keys found in environment variables
#> ℹ Active account: "default" (the default)
#>
#> ── API key
#> ✖ No API key found for account "default"
#>
#> ── Connectivity
#> ✔ Internet: online
#> ℹ Skipping API connection test
#>
#> ── Recommendations
#> • Create an API key at <https://tally.so/settings/api-keys>
#> • Set `TALLY_API_KEY` in your '.Renviron' (e.g. with
#> `usethis::edit_r_environ()`)
```

tallyr couldn't find a key for the active account. To fix it, follow the
recommendations: create a key at [Tally Settings > API
keys](https://tally.so/settings/api-keys) and set it in your `.Renviron`.

Two common traps when the key *is* in `.Renviron` but still isn't found:

- **R hasn't been restarted.** `.Renviron` is only read at startup, so
  restart R after editing it.
- **The key is set for a different account.** tallyr looks in
  `TALLY_API_KEY` for the default account, and in `TALLY_API_KEY_<NAME>`
  for the named account `<NAME>`. If the Accounts section shows a key
  available under another name, the report will suggest switching:

```{r, eval=FALSE}
#> ── Recommendations
#> • Create an API key at <https://tally.so/settings/api-keys>
#> • Set `TALLY_API_KEY` in your '.Renviron' (e.g. with
#> `usethis::edit_r_environ()`)
#> • Or switch to an account with a key: `tally_use_account("work")`
```

See `vignette("multiple-accounts")` for how named accounts work.

## Key found, but rejected by the API

```{r, eval = FALSE}
tally_sitrep()
#>
#> ── tallyr situation report ─────────────────────────────────────────────
#>
#> ── Accounts
#> ℹ Account with a key available: "default"
#> ℹ Active account: "default" (the default)
#>
#> ── API key
#> ✔ API key found
#> ℹ Source: Environment variable: TALLY_API_KEY
#> ✔ Key format looks valid (starts with "tly-")
#>
#> ── Connectivity
#> ✔ Internet: online
#> ✖ Tally API connection failed: HTTP 401 Unauthorized.
#>
#> ── Recommendations
#> • Check your key is current, or regenerate it at
#> <https://tally.so/settings/api-keys>
```

A key was found and sent, but Tally refused it -- typically because the
key was revoked or mistyped. Compare the value in your `.Renviron`
against the keys listed at [Tally Settings > API
keys](https://tally.so/settings/api-keys), or generate a fresh one.

Relatedly, Tally API keys start with `tly-`. If yours doesn't, the API
key section warns that it may not be a Tally key at all -- a hint that
the wrong value (a key for some other service, say) ended up in the
variable:

```{r, eval=FALSE}
#> ── API key
#> ✔ API key found
#> ℹ Source: Environment variable: TALLY_API_KEY
#> ! Key does not start with "tly-", so may not be a Tally API key
```

## Offline

```{r, eval=FALSE}
#> ── Connectivity
#> ✖ Internet: offline
#> ℹ Skipping API connection test
#>
#> ── Recommendations
#> • Reconnect to the internet to use the Tally API
```

Nothing wrong with your tallyr setup -- the API just can't be reached.

## Where tallyr looks for keys

For reference, when a tallyr function needs a key for the **default**
account, it checks, in order:

1. The `tallyr.api_key` R option (useful for setting a key for a single
   session, with `options(tallyr.api_key = "tly-...")`).
1. The `TALLY_API_KEY` environment variable.

A **named** account's key comes only from its `TALLY_API_KEY_<NAME>`
environment variable; the option is never consulted. The Source line in
`tally_sitrep()`'s API key section tells you which of these supplied the
active key, which helps when a stale option or an unexpected environment
variable is shadowing the key you meant to use.

If you can't tell which account is even active, the Accounts section
reports it, along with how it was chosen: the default, set by
`tally_use_account()`, or passed as an `account` argument.
